Online Laterality Battery Pre-Processing - short version

### 2020-07-19 battery-preprocessing_v6_DB.Rmd
Some more tweaks to z-tests. All polarity set so L hemisphere advantage is positive, to be consistent with main papers in the literature - and with Doppler.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
library(beeswarm) #just a kind of plot Dorothy likes
```

@@

@@

```{r domakeLI}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_CF_RT_LI"
polarity <-1

@@

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-1
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

```{r domakeLidich}
#We add Dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_DL_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_DL_RT_LI"
polarity <-1

@@

origdat<-DL_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_DL_acc_LI"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_DL_RT_LI"
polarity <-1
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

```{r domakeZ_dich}
#We add dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_Dich_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_Dich_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

# accuracy
origdat$X1 <- origdat$p.corr
origdat$side <- as.factor(origdat$side)
levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
mycolname<-"Day1.RDT_p.corr_LI"
if(d==2){mycolname<-"Day2.RDT_p.corr_LI"}
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# RT
origdat$X1 <- origdat$RT
mycolname<-"Day1.RDT_RT_LI"
if(d==2){mycolname<-"Day2.RDT_RT_LI"}
polarity <- (1)
mycutoff <- 0 #no cutoff for logRT data
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}

@@

  myfile<-filter(myfile,side %in% c('left','right'))

  nsubs <- nrow(allsum)
  allsum$tlat_RDT<-NA

  for (i in 1:nsubs) {
    w<-which(myfile$subject==allsum$subject[i])
    if(length(w)>10)
    {
      myt<-t.test(myfile$RT[w]~myfile$side[w]) #this is based on logRT
      allsum$tlat_RDT[i]<-myt$statistic
    }
  }
  if (d==1){
    maxcol <-length(colnames(allsum))
    colnames(allsum)[maxcol]<-c('tlat_RDT_1')
  }
  if (d==2){
    maxcol <-length(colnames(allsum))
    colnames(allsum)[maxcol]<-c('tlat_RDT_2')
  }
}

colcount<-length(colnames(allsum))

w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
allsum[w,(colcount-1):colcount]<-NA
w<-which(is.na(allsum$Day2.RDT_RT_LI))
allsum[w,(colcount-1):colcount]<-NA

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# libraries 
library(readr)
library(yarrr)
library(dplyr)
library(tidyr)
library(car)
library(ggplot2)
library(knitr)
require(readr)  # for read_csv()
require(dplyr)  # for mutate()
require(purrr)  # for map(), reduce()
library(GGally)
require(stringr) # string matching
library(ggpirate)
library(corrr) #easier correlations!
library(beeswarm) #just a kind of plot Dorothy likes
library(BSDA) #for ztest
```

@@

# 19 Jul 2020:settled on setting polarity so that L hemisphere advantage is positive

@@

```{r domakeLI}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Chimeric faces'
origdat<-chim_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_CF_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_CF_RT_LI"
polarity <-(-1)

@@

origdat<-chim_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_CF_acc_LI"
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_CF_RT_LI"
polarity <-(-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

```{r domakeLidich}
#We add Dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$accurate
mycolname<-"Day1_DL_acc_LI"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# now do RT
origdat$X1 <- origdat$RT
mycolname<-"Day1_DL_RT_LI"
polarity <-(-1)

@@

origdat<-DL_dat2
origdat$X1 <- origdat$accurate
mycolname<-"Day2_DL_acc_LI"
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day2_DL_RT_LI"
polarity <-(-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

```{r domakeZ_dich}
#We add dichotic LIs to allsum for day 1 and then day2
task<-'Dichotic listening'
origdat<-DL_dat1
origdat$X1 <- origdat$N
mycolname<-"Day1_Dich_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-DL_dat2
origdat$X1 <- origdat$N
mycolname<-"Day2_Dich_acc_Z"
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

@@

# accuracy
origdat$X1 <- origdat$p.corr
origdat$side <- as.factor(origdat$side)
levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
mycolname<-"Day1.RDT_p.corr_LI"
if(d==2){mycolname<-"Day2.RDT_p.corr_LI"}
polarity<- (-1) #so LI is positive in predicted direction
mycutoff <- 75
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

# RT
origdat$X1 <- origdat$RT
mycolname<-"Day1.RDT_RT_LI"
if(d==2){mycolname<-"Day2.RDT_RT_LI"}
polarity <- (-1)
mycutoff <- 0 #no cutoff for logRT data
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}

@@

  nsubs <- nrow(allsum)
  allsum$tempt<-NA #initialise a new column

  for (i in 1:nsubs) {
    mysub<-filter(myfile,side %in% c('left','right'),subject==allsum$subject[i])
    if(nrow(mysub)>10)
    {
      myt<-t.test(mysub$RT~mysub$side) #this is based on logRT
      allsum$tempt[i]<-myt$statistic
    }
  }
  wc<-which(colnames(allsum)=='tempt')
  colnames(allsum)[wc]<-paste0('tlat_RDT_',d)

w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
allsum$tlat_RDT_1[w]<-NA
w<-which(is.na(allsum$Day2.RDT_RT_LI))
allsum$tlat_RDT_2[w]<-NA

@@

In fact, for better consistency with the other measures, could do z-test instead of t-test. These converge to the same thing with samples > around 30, which we have here.  So here is that calculation.

```{r make.zlatRDT}
for (d in 1:2){
  myfile<-RDT_dat1
  if(d==2){
    myfile<-RDT_dat2}

  nsubs <- nrow(allsum)
  allsum$tempt<-NA #initialise a new column

  for (i in 1:nsubs) {
    mysub<-filter(myfile,side %in% c('left','right'),subject==allsum$subject[i])
    if(nrow(mysub)>10)
    {

      mymean<-aggregate(RT~side,mysub,FUN = mean)
      mysd<-aggregate(RT~side,mysub,FUN = sd)
        myN<-aggregate(RT~side,mysub,FUN = length)
      se1<-mysd[1,2]/sqrt(myN[1,2])
      se2<-mysd[2,2]/sqrt(myN[2,2])
      se<-sqrt(se1^2+se2^2)
      allsum$tempt[i]<-(mymean[1,2]-mymean[2,2])/se
    }
  }
  wc<-which(colnames(allsum)=='tempt')
  colnames(allsum)[wc]<-paste0('zlat_RDT_',d)

w<-which(is.na(allsum$Day1.RDT_RT_LI)) #remove those with NA in LI
allsum$zlat_RDT_1[w]<-NA
w<-which(is.na(allsum$Day2.RDT_RT_LI))
allsum$zlat_RDT_2[w]<-NA

#compare t-value from day 1 and day 2 
myr <- cor(allsum$zlat_RDT_1,allsum$zlat_RDT_2,use='complete.obs')
plot(allsum$zlat_RDT_1,allsum$zlat_RDT_2,xlim=c(-5,5),ylim=c(-5,5))
abline(v=0)
abline(h=0)
abline(v=-1.65,lty=2,col='red') 
abline(h=-1.65,lty=2,col='red')
abline(v=1.65,lty=2,col='red') 
abline(h=1.65,lty=2,col='red')
text(-3,3,paste0('r = ',round(myr,3)))

plot(allsum$Day1.RDT_RT_LI,allsum$zlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
plot(allsum$tlat_RDT_1,allsum$zlat_RDT_1,xlim=c(-1.5,1.5),ylim=c(-5,5))
abline(0,1)

#Confirms tlat is same as zlat for these sample sizes!
```
